Borders and Gateways: Measuring and Analyzing National AS Chokepoints

### 2018-08-01 Paper/chokepoint.tex
Spellcheck run through

We define a BGP path of length $n$ from source AS $a_0$ to destination AS $a_n$
as a sequence $p_{1\rightsquigarrow n} = \{a_i : i \in [1,n]\}$.  We define a
\textit{routing tree} to be a tree rooted at a single destination AS and
composed of every acceptable BGP path from the various source ASes that can
reach the destination AS. Note that if multiple paths exist from a source
to a destination, the routing tree is technicailly a DAG, but we use the traditional convention and call
all such data structures routing trees. The
chokepoint potential of a border AS is the ratio of the number of paths in (or out of) the country that contain the border AS to
the total number of paths in (or out of) the same country.
%intercepted by all border ASes belonging to
%the same country.
Formally, for country $c$, let $P = \{p_{i \rightsquigarrow
j} : g(a_i)=c \text{ and } g(a_j)\neq c\} $, i.e.\ the set of all paths
originating in $c$ and ending outside $c$. Then the chokepoint potential,
$cp(\cdot)$, of a border AS $a_u$ with $g(a_u) = c$ is
\begin{equation}
  \label{eq:chokepoint}
  cp(a_u) = \frac{|\{p: p \in P \text{ and } a_u \in P\}|}{|P|}
\end{equation}

The above definition applies to paths originating in country
$c$ and routing to a different country, which we will refer to as
\textit{outgoing} chokepoint potential or $cpo$. We similarily define
\textit{incoming} chokepoint potential or $cpi$, by redefining $P$ as $\{p_{i
\rightsquigarrow j}: g(a_i) \neq c \text{ and } g(a_j) = c\}$. That is all
paths originating outside of $c$ but terminating in $c$ and recomputing
Equation~\ref{eq:chokepoint}.

>>>>

We define a BGP path of length $n$ from source AS $a_0$ to destination AS $a_n$
as a sequence $p_{1\rightsquigarrow n} = \{a_i : i \in [1,n]\}$.  We define a
\textit{routing tree} to be a tree rooted at a single destination AS and
composed of every acceptable BGP path from the various source ASes that can
reach the destination AS. Note that if multiple paths exist from a source
to a destination, the routing tree is technically a DAG, but we use the traditional convention and call
all such data structures routing trees. The
chokepoint potential of a border AS is the ratio of the number of paths in (or out of) the country that contain the border AS to
the total number of paths in (or out of) the same country.
%intercepted by all border ASes belonging to
%the same country.
Formally, for country $c$, let $P = \{p_{i \rightsquigarrow
j} : g(a_i)=c \text{ and } g(a_j)\neq c\} $, i.e.\ the set of all paths
originating in $c$ and ending outside $c$. Then the chokepoint potential,
$cp(\cdot)$, of a border AS $a_u$ with $g(a_u) = c$ is
\begin{equation}
  \label{eq:chokepoint}
  cp(a_u) = \frac{|\{p: p \in P \text{ and } a_u \in P\}|}{|P|}
\end{equation}

The above definition applies to paths originating in country
$c$ and routing to a different country, which we will refer to as
\textit{outgoing} chokepoint potential or $cpo$. We similarly define
\textit{incoming} chokepoint potential or $cpi$, by redefining $P$ as $\{p_{i
\rightsquigarrow j}: g(a_i) \neq c \text{ and } g(a_j) = c\}$. That is all
paths originating outside of $c$ but terminating in $c$ and recomputing
Equation~\ref{eq:chokepoint}.

### 2018-08-01 Paper/conclusion.tex
Spellcheck run through

% Points of Presence
\subsection{Other Chokepoint Measures}
There are several ways that a country can create chokepoints, either intentionally or accidentally: taking advantage of Internet Exchange Points (IXPs),
limiting the number of physical connections that cross the border, centralizing
the DNS infrastructure within the country, or working
with---and supporting---a set of ISPs that have international connections.
Each of these combines two basic elements: limiting the number of organizations that the government must interact with and creating physical locations where traffic can be controlled.
We chose to define chokepoint potential in terms of the AS graph instead of physical locations because the AS graph captures the most interesting physical attributes of the Internet, and
virtualization of the physical and datalink layers on the Internet make physical locations less meaningful. A single undersea cable can be multiplexed to serve as several different links in
the AS graph, and ASes in one country can have physical Points of Presence (PoPs) in several other countries (e.g., China Telecom in Pasadena, California). Thus, the AS graph is a good approximation
of overall chokepoint potential, and information such as geography and the physical locations of routers does not add meaningfully to the analysis without considering every individual piece of Internet
infrastructue on a case-by-case basis. Additionally, the dramatic shifts in chokepoint potential over time depicted in our analysis are unrelated to constant geography, so our AS-level study is important
to show time evolution of chokepoints.

@@

\subsection{Future Work}
% OONI data
While linking chokepoint potential to FOTN scores is a substantial contribution, it still stands that FOTN can serve only as a proxy for censorship.
This is useful for large trends and general understanding, but a more fine-grained approach could be used to interpret the direct connection between
actual censorship events and the shifts in the AS graph. The Open Observatory of Network Interference, or OONI, \cite{OONI} provides Internet users around
the world with a tool called the ooniprobe. The ooniprobe lets users run a suite of tests to identify censorship anamolies of various types, and the results are
recorded in the large OONI database. Matching up changes in OONI measurements, such as increased censorship campaigning in an authoritarian nation, with shifts
in chokepoint potential would be a major step in understanding the interplay of censorship and AS-level chokepoints. This process involves designing a way to classify
censorship events and chokepoint potential changes, and as such lies beyond the scope of this study.

\section{Conclusion}
% Chokepoint Potential
This paper addresses the question of how Internet structure relates to 
international boundaries and how it is changing over time.  By defining chokepoint
potential, both for individual border ASes and aggregated by country,
we provide a quantitative way to measure these effects in terms of the ease
with which a government can control routing paths across its borders. 
Our technique for generating BGP paths and evaluating chokepoints
using \toolname{} provides public domain software and routing trees for the entire AS graph.
%a monitoring capability of BGP dynamics that was previously not possible.
We have taken advantage of efficient simulation, standard AS relationship datasets, and cross-platform
design principles so that this tool will be readily deployable for future research.
% Ecolution over time
\par We used \toolname{} to investigate how the global AS graph has
evolved over time, finding interesting trends and outliers. These
studies reveal a tendency in large nations to concentrate paths across
their borders into a small number of ASes, despite the increasing
connectivity and size of the AS graph. Additionally, we find that border
chokepoints in some nations known to conduct censorship and
surveillance have steadily increased in strength.
% Internet Freedom
\par
We studied the relationship between NCP and two evaluations of openness, Freedom of the Net  and Freedom of the press, finding stastistically significant relationships in both cases.
%These results suggest that it is worthwhile to consider border AS chokepoints when studying
This relationship implies that when censorship and surveillance are
being evaluated at the national level, it is worthwhile to consider
the border AS chokepoints that are resident to the nation in question.

>>>>

% Points of Presence
\subsection{Other Chokepoint Measures}
There are several ways that a country can create chokepoints, either intentionally or accidentally: taking advantage of Internet Exchange Points (IXPs),
limiting the number of physical connections that cross the border, centralizing
the DNS infrastructure within the country, or working
with---and supporting---a set of ISPs that have international connections.
Each of these combines two basic elements: limiting the number of organizations that the government must interact with and creating physical locations where traffic can be controlled.
We chose to define chokepoint potential in terms of the AS graph instead of physical locations because the AS graph captures the most interesting physical attributes of the Internet, and
virtualization of the physical and datalink layers on the Internet make physical locations less meaningful. A single undersea cable can be multiplexed to serve as several different links in
the AS graph, and ASes in one country can have physical Points of Presence (PoPs) in several other countries (e.g., China Telecom in Pasadena, California). Thus, the AS graph is a good approximation
of overall chokepoint potential, and information such as geography and the physical locations of routers does not add meaningfully to the analysis without considering every individual piece of Internet
infrastructure on a case-by-case basis. Additionally, the dramatic shifts in chokepoint potential over time depicted in our analysis are unrelated to constant geography, so our AS-level study is important
to show time evolution of chokepoints.

@@

\subsection{Future Work}
% OONI data
While linking chokepoint potential to FOTN scores is a substantial contribution, it still stands that FOTN can serve only as a proxy for censorship.
This is useful for large trends and general understanding, but a more fine-grained approach could be used to interpret the direct connection between
actual censorship events and the shifts in the AS graph. The Open Observatory of Network Interference, or OONI, \cite{OONI} provides Internet users around
the world with a tool called the ooniprobe. The ooniprobe lets users run a suite of tests to identify censorship anomalies of various types, and the results are
recorded in the large OONI database. Matching up changes in OONI measurements, such as increased censorship campaigning in an authoritarian nation, with shifts
in chokepoint potential would be a major step in understanding the interplay of censorship and AS-level chokepoints. This process involves designing a way to classify
censorship events and chokepoint potential changes, and as such lies beyond the scope of this study.

\section{Conclusion}
% Chokepoint Potential
This paper addresses the question of how Internet structure relates to 
international boundaries and how it is changing over time.  By defining chokepoint
potential, both for individual border ASes and aggregated by country,
we provide a quantitative way to measure these effects in terms of the ease
with which a government can control routing paths across its borders. 
Our technique for generating BGP paths and evaluating chokepoints
using \toolname{} provides public domain software and routing trees for the entire AS graph.
%a monitoring capability of BGP dynamics that was previously not possible.
We have taken advantage of efficient simulation, standard AS relationship datasets, and cross-platform
design principles so that this tool will be readily deployable for future research.
% Ecolution over time
\par We used \toolname{} to investigate how the global AS graph has
evolved over time, finding interesting trends and outliers. These
studies reveal a tendency in large nations to concentrate paths across
their borders into a small number of ASes, despite the increasing
connectivity and size of the AS graph. Additionally, we find that border
chokepoints in some nations known to conduct censorship and
surveillance have steadily increased in strength.
% Internet Freedom
\par
We studied the relationship between NCP and two evaluations of openness, Freedom of the Net  and Freedom of the press, finding statistically significant relationships in both cases.
%These results suggest that it is worthwhile to consider border AS chokepoints when studying
This relationship implies that when censorship and surveillance are
being evaluated at the national level, it is worthwhile to consider
the border AS chokepoints that are resident to the nation in question.

### 2018-08-01 Paper/intro.tex
Spellcheck run through

We developed a
suite of tools, called BGP Simulation, Analysis, and Storage (\toolname{}), for studying national 
chokepoints on
the AS graph efficiently. To illustrate these ideas and tools, we
study how NCP correlates with two independent
measures of civil liberty, finding that a significant relationship exists between NCP and each measure.
The paper extends earlier research on AS topology in several ways: it introduces chokepoint potential, it presents open-source cross-platform tools and datasetes for simulating Border Gateway Protocol (BGP) paths,
determining NCP, and analyzing changes over time.  In addition, we report and analyze data for several countries of interest.
%Our measure of border ASes, chokepoint potential, defined in detail
%in section 3, is a succinct way to estimate the important properties of
%border ASes related to the ability of a nation to censor and surveil, and our aggregate extension of this measure to
%the national level, national breakthrough potential, is an effective way to evaluate and
%compare nations and analyze the evolution of the AS-level censorship and surveillance landscape.\Ben[inline]{This paragraph seems a little redundant now. It is a combo of other previous writing, but I am not sure what to cut.}

@@

% Rest of Paper Layout
\par
The remainder of the paper is organized as follows: Section II provides relevent background for
the problems we are investigating; Section III introduces and defines the measures of chokepoint
potential and national chokepoint potential; Section IV describes \toolname{} for simulating and evaluating 
BGP networks; Section V
explains the experimental setup and data sources; Sections VI, VII, and VIII contain experimental results, related work, and discussion,
respectively.  Section IX concludes the paper.

>>>>

We developed a
suite of tools, called BGP Simulation, Analysis, and Storage (\toolname{}), for studying national 
chokepoints on
the AS graph efficiently. To illustrate these ideas and tools, we
study how NCP correlates with two independent
measures of civil liberty, finding that a significant relationship exists between NCP and each measure.
The paper extends earlier research on AS topology in several ways: it introduces chokepoint potential, it presents open-source cross-platform tools and datasets for simulating Border Gateway Protocol (BGP) paths,
determining NCP, and analyzing changes over time.  In addition, we report and analyze data for several countries of interest.
%Our measure of border ASes, chokepoint potential, defined in detail
%in section 3, is a succinct way to estimate the important properties of
%border ASes related to the ability of a nation to censor and surveil, and our aggregate extension of this measure to
%the national level, national breakthrough potential, is an effective way to evaluate and
%compare nations and analyze the evolution of the AS-level censorship and surveillance landscape.\Ben[inline]{This paragraph seems a little redundant now. It is a combo of other previous writing, but I am not sure what to cut.}

@@

% Rest of Paper Layout
\par
The remainder of the paper is organized as follows: Section II provides relevant background for
the problems we are investigating; Section III introduces and defines the measures of chokepoint
potential and national chokepoint potential; Section IV describes \toolname{} for simulating and evaluating 
BGP networks; Section V
explains the experimental setup and data sources; Sections VI, VII, and VIII contain experimental results, related work, and discussion,
respectively.  Section IX concludes the paper.

### 2018-08-01 Paper/results.tex
Spellcheck run through

For each year there appear to be two general modes.  Nations that are
not free or partly free tend to have high NCP, while free nations have
variable NCPs. There are interesting outliers for both situations,
however. Countries like Estonia (EE) or Iceland (IS) are very free but
require few border ASes to control most of their paths. This is likely
because their overall AS counts are low. As we saw earlier, Russia
(RU) is considered not free by FOTN, but requires a large number of
ASes to control most of its paths. This suggests that censorship
in Russia is implemented without AS-level
chokepoints.

>>>>

The trend holds for all years investigated that countries that are
not free or partly free tend to have high NCP, while free nations have
variable NCPs. There are interesting outliers for both situations,
however. Countries like Estonia and Iceland are very free but
require few border ASes to control most of their paths. This is likely
because their overall AS counts are low. As we saw earlier, Russia
is considered not free by FOTN, but requires a large number of
ASes to control most of its paths. This suggests that censorship
in Russia is implemented without AS-level
chokepoints.

### 2018-08-01 Paper/tool.tex
Minor trim in tool section
Spellcheck run through

% Preprocessing and Data pipeline
\toolname{} determines a set of plausible BGP paths through simulation and calculates AS
chokepoint potential and national chokepoint potential.  \toolname{} takes as input AS relationship data, such as that provided by
CAIDA~\cite{CAIDA}.  It also takes as input a set of AS
country codes (that identifies which nation an AS belongs to) for
determining chokepoint potential.  \toolname{}
uses an algorithm based on BGPSim \cite{quicksand} to
generate a set of routing trees. \toolname{} 
calclates both AS and national chokepoint potential for any country.

Our experiments used country codes returned from Team Cymru's IP to
ASN whois service \cite{cymru} to determine which ASes were registered
to which nation.  This service maintains only the most recent
registration of each AS.  To control for this possible source of
error, we retain ASes in the path simulation, even if they were
registered more recently than the current timestamp being studied. This is because
the relationships of these ASes and paths through these ASes are still valid.
We do not assign these ASes chokepoint potentials, as it is not clear which nation
these ASes are registered for. We took this approach
because the datasets of AS registrations for the various ASN
registries tend to have more missing data at specific timestamps.
We do not expect this to have a considerable effect on our results
because few ASes do not have a definite nation (in the best case in the most recent
test this is less than 1\% of ASes, in the worst case in the earliest test this is around 12\% of ASes.)
Additionally, not all of the ASes that are unlinked to a nation are border ASes, and those will not
affect our results at all. Furthermore, it is unlikely that this small portion of ASes will be a majority
of border ASes for any single country, so the country level experiments are unlikely to suffer from this
hole in the data.

@@

% Chokepoint calculation
\par Once \toolname{} generates the routing trees, they can be processed to
determine chokepoint potentials or by other researchers interested in other questions. To calculate chokepoint potential, \toolname{} iterates over every path
between each AS-pair. Because we use the same random tie-break method as BGPSim,this process returns 
exactly one path between each AS-pair considered, even if multiple options exist in the
routing tree. Tiebreaks do not have a noticable effect on our results. For instance, the maximum
standard deviation for the chokepoint potential of any AS in 5 runs of one of our timestamps was 
$\leq$ 0.005, meaning the chokepoint potentials for each AS barely changed depending on the tiebreak
decisions.
Once the path has been determined, it is traversed to identify border nodes and increment their counts, both
for outgoing and incoming paths.  Thus, only one traversal is conducted
per path. Additionally, the number of paths of each type that belong to each
nation is tallied. \toolname{} takes the resulting chokepoint
potentials and generates national chokepoint potentials for each nation.

Calculating routing trees for the entire AS graph takes time proportional to the $|V|^2$ where $|V|$ is the total number of ASes; calculating chokepoint potential takes $|V|^2D$ where $D$ is the average depth of a routing tree. For example,
it took about 12 wall clock hours to compute all of the routing trees for the Jan. 2017 dataset, running on 4 Intel Xeon E5-2680 V4 CPUs.  For each dataset, this is a one-time cost because the routing trees are stored externally.  [URL blineded for peer review]

>>>>

% Preprocessing and Data pipeline
\toolname{} determines a set of plausible BGP paths through simulation and calculates AS
chokepoint potential and national chokepoint potential.  \toolname{} takes as input AS relationship data, such as that provided by
CAIDA~\cite{CAIDA}.  It also takes as input a set of AS
country codes (that identifies which nation an AS belongs to) for
determining chokepoint potential.  \toolname{}
uses an algorithm based on BGPSim \cite{quicksand} to
generate a set of routing trees. \toolname{} 
calculates both AS and national chokepoint potential for any country.

Our experiments used country codes returned from Team Cymru's IP to
ASN whois service \cite{cymru} to determine which ASes were registered
to which nation.  This service maintains only the most recent
registration of each AS.  To control for this possible source of
error, we retain ASes in the path simulation, even if they were
registered more recently than the current timestamp being studied. This is because
the relationships of these ASes and paths through these ASes are still valid.
We do not assign these ASes chokepoint potentials, as it is not clear which nation
these ASes are registered for. We took this approach
because the datasets of AS registrations for the various ASN
registries tend to have more missing data at specific timestamps.
We do not expect this to have a considerable effect on our results
because few ASes do not have a definite nation (in the best case in the most recent
test this is less than 1\% of ASes, in the worst case in the earliest test this is around 12\% of ASes.)
Additionally, not all of the ASes that are unlinked to a nation are border ASes, and those will not
affect our results at all.

@@

% Chokepoint calculation
\par Once \toolname{} generates the routing trees, they can be processed to
determine chokepoint potentials or by other researchers interested in other questions. To calculate chokepoint potential, \toolname{} iterates over every path
between each AS-pair. Because we use the same random tie-break method as BGPSim,this process returns 
exactly one path between each AS-pair considered, even if multiple options exist in the
routing tree. Tiebreaks do not have a noticeable effect on our results. For instance, the maximum
standard deviation for the chokepoint potential of any AS in 5 runs of one of our timestamps was 
$\leq$ 0.005, meaning the chokepoint potentials for each AS barely changed depending on the tiebreak
decisions.
Once the path has been determined, it is traversed to identify border nodes and increment their counts, both
for outgoing and incoming paths.  Thus, only one traversal is conducted
per path. Additionally, the number of paths of each type that belong to each
nation is tallied. \toolname{} takes the resulting chokepoint
potentials and generates national chokepoint potentials for each nation.

Calculating routing trees for the entire AS graph takes time proportional to the $|V|^2$ where $|V|$ is the total number of ASes; calculating chokepoint potential takes $|V|^2D$ where $D$ is the average depth of a routing tree. For example,
it took about 12 wall clock hours to compute all of the routing trees for the Jan. 2017 dataset, running on 4 Intel Xeon E5-2680 V4 CPUs.  For each dataset, this is a one-time cost because the routing trees are stored externally.  [URL blinded for peer review]
